# STAT209 HW3 

Seokjun Choi


## Problem 1

*Code blocks are R scripts of this problem.*

*Executable stand-alone R script: https://github.com/letsjdosth/GLM/blob/main/hw3q1.r*

I will use the same notation given in the problem.

### 1-(a)

Let me fit three models with different links, logit, probit, and complementary log-log.
It is easy because base-R's 'glm' and 'binomial' functions offer everything I need.

```{r}
log_dose <- c(1.6907, 1.7242, 1.7552, 1.7842, 1.8113, 1.8369, 1.8610, 1.8839)
beetles <- c(59, 60, 62, 56, 63, 59, 62, 60) #m
killed <- c(6, 13, 18, 28, 52, 53, 61, 60) #y

#1a
logit_fit = glm(cbind(killed, beetles-killed)~log_dose, family=binomial(link="logit"))
summary(logit_fit)
probit_fit = glm(cbind(killed, beetles-killed)~log_dose, family=binomial(link="probit"))
summary(probit_fit)
cloglog_fit = glm(cbind(killed, beetles-killed)~log_dose, family=binomial(link="cloglog"))
summary(cloglog_fit)
```

Getting deviance residual is also easy in R.
All functions that we need are already implemented in base-R.

```{r}
res_logit = residuals(logit_fit, "deviance")
res_probit = residuals(probit_fit, "deviance")
res_cloglog = residuals(cloglog_fit, "deviance")

plot(log_dose, res_logit, col="orange", xlim=c(1.65, 1.9), cex=3)
points(log_dose, res_probit, col="red", cex=3)
points(log_dose, res_cloglog, col="blue", cex=3)
abline(h=0)
```

The above plot is an overlaid residual plot.
I put three models' residuals in the plot.
Orange dots are from the logit model, red dots are from the probit model, and blue dots are from the cloglog model.

Logit and probit fits are similar.
Comparing them to the cloglog model, we may say that the cloglog model performs better than the others,
generally having smaller residuals in view of abstract values (norm values).

Next, I show the $\pi_i$ fits versus the covariate.

```{r}
inv_logit <- function(x, beta_0, beta_1){
    return(exp(beta_0 + beta_1*x) / (1 + exp(beta_0 + beta_1*x)))
}
inv_probit <- function(x, beta_0, beta_1){
    return(pnorm(beta_0 + beta_1*x))
}
inv_cloglog <- function(x, beta_0, beta_1){
    return(1-exp(-exp(beta_0 + beta_1*x)))
}
# grid = seq(1.65, 1.9, 0.005)
plot(log_dose, killed/beetles, xlim=c(1.65, 1.9), cex=3)
curve(inv_logit(x, logit_fit$coefficients[1], logit_fit$coefficients[2]), add=T, col="orange")
curve(inv_probit(x, probit_fit$coefficients[1], probit_fit$coefficients[2]), add=T, col="red")
curve(inv_cloglog(x, cloglog_fit$coefficients[1], cloglog_fit$coefficients[2]), add=T, col="blue")
```

Again, the logit model fit has an orange color, the probit model is red, and the cloglog model uses blue.
Black dots are observed ratio, calculated by $y_i/m_i$.

We can see that the cloglog fit(blue) is closer to the observed ratio for almost all points.
The asymmetry of cloglog link may help to fit better.

### 1-(b)

Please see the attached note separately.

### 1-(c)

Cause all MLE values are given, we can track back fitted $\pi_i$ and the model's likelihood value.
Because equation (1.1) directly gives us the inverse link function, we can use it to get fitted $\pi_i$.

```{r}
#1c: alpha-power logit link
pll_beta0 = -113.625
pll_beta1 = 62.5
pll_alpha = 0.279

pll_inv <- function(x, beta_0, beta_1, alpha){
    nu = beta_0 + beta_1*x
    return(exp(alpha*nu)/((1+exp(nu))^alpha))
}
(pll_fit = pll_inv(log_dose, pll_beta0, pll_beta1, pll_alpha))
```

The above output shows the fitted $\pi_i$s at the $i$th data point.

Plus, let me put the $\pi$ fits of the powered-logit link model over the last plot I showed.

```{r}
plot(log_dose, killed/beetles, xlim=c(1.65, 1.9), cex=3)
curve(inv_logit(x, logit_fit$coefficients[1], logit_fit$coefficients[2]), add=T, col="orange")
curve(inv_probit(x, probit_fit$coefficients[1], probit_fit$coefficients[2]), add=T, col="red")
curve(inv_cloglog(x, cloglog_fit$coefficients[1], cloglog_fit$coefficients[2]), add=T, col="blue")
curve(pll_inv(x, pll_beta0, pll_beta1, pll_alpha), add=T, col="green")
```

The green curve is the new powered-logit link model fit.
In the left tail, it assigns more $pi$ than the cloglog fit.
In the middle area (between $1.75$ and $1.80$,) the powered-logit fit is slightly lower $\pi$ than the cloglog fit,
making a closer fit to $y/m$.
In the right tail, the cloglog and the powered logit link fits are very similar.

The powered logit fit is quite good, carrying a very flexible asymmetry.
But it is not very different from the cloglog fit, though.
If we want a simpler model, the cloglog fit may be better.


```{r}
pll_dev_res <- function(m, y, fitted_pi){
    red_theta = log(fitted_pi/(1-fitted_pi))
    full_theta = log(y/(m-y))
    d = 2*(y*(full_theta - red_theta) + m*log((1+exp(red_theta))/(1+exp(full_theta))))
    sign = ifelse((y/m - fitted_pi)>0, +1, -1)
    return(sign*sqrt(d))
}

#pll dev.resid
pll_dev_res(beetles, killed, pll_fit) #cause m=y for last point, it gives NaN
```

Finally, the last line's output gives us the residual values.
The function 'pll_dev_res' calculates the deviance residual of the new model.
Since $m_i=y_i$ at $i=8$, my poor, non-adjusted function yields 'NaN'.
We may use the adjusted deviance residual formula for $y=m$ or $y=0$ cases, but let me skip it here.

Other values, except for one, are quite close to 0.

### 1-(d)

Below are AIC and BIC values for each model.

```{r}
# 1d. AIC / BIC
pll_log_lik = 0
for(i in 1:8){
    pll_log_lik = pll_log_lik + dbinom(killed[i],beetles[i],pll_fit[i], log=TRUE)
}
# (pll_log_lik)

pll_num_param = 3
(pll_AIC = -2*pll_log_lik + 2*pll_num_param)
(pll_BIC = -2*pll_log_lik + pll_num_param * log(8))

AIC(logit_fit)
BIC(logit_fit)
AIC(probit_fit)
BIC(probit_fit)
AIC(cloglog_fit)
BIC(cloglog_fit)

```

Interestingly, both AIC and BIC select the cloglog fit.
It consists of what we found earlier.

The basic logit and probit fit were worse than the cloglog or powered logit fit. So, they also have higher AIC and BIC values.

Since the powered logit model has one more parameter, it gets more penalty from AIC and BIC calculations.
But the fit of the powered logit link model was not much improved from the cloglog fit, according to the previous observation.
That's why AIC and BIC of powered logit link fit are higher than the cloglog fit.