# STAT209 HW3 

Seokjun Choi


## Problem 1

*Code blocks are R scripts of this problem.*

*Executable stand-alone R script: https://github.com/letsjdosth/GLM/blob/main/hw3q1.r*

I will use the same notation given in the problem.

### 1-(a)

Let me fit three models having different links, logit, probit, and complementary log-log, respectively.
It is easy cause base-R's 'glm' and 'binomial' functions offer everything that I need.

```{r}
log_dose <- c(1.6907, 1.7242, 1.7552, 1.7842, 1.8113, 1.8369, 1.8610, 1.8839)
beetles <- c(59, 60, 62, 56, 63, 59, 62, 60) #m
killed <- c(6, 13, 18, 28, 52, 53, 61, 60) #y

#1a
logit_fit = glm(cbind(killed, beetles-killed)~log_dose, family=binomial(link="logit"))
summary(logit_fit)
probit_fit = glm(cbind(killed, beetles-killed)~log_dose, family=binomial(link="probit"))
summary(probit_fit)
cloglog_fit = glm(cbind(killed, beetles-killed)~log_dose, family=binomial(link="cloglog"))
summary(cloglog_fit)
```

Getting deviance residual is also easy in R.
All functions that we need are already implemented in base-R.

```{r}
res_logit = residuals(logit_fit, "deviance")
res_probit = residuals(probit_fit, "deviance")
res_cloglog = residuals(cloglog_fit, "deviance")

plot(log_dose, res_logit, col="orange", xlim=c(1.65, 1.9), cex=3)
points(log_dose, res_probit, col="red", cex=3)
points(log_dose, res_cloglog, col="blue", cex=3)
abline(h=0)
```

The above plot is a overlaid residual plot.
I put three model's residuals in the plot.
Orange dots are of the logit model's, red dots are of the probit model's, and blue dots are from the cloglog model.

It seems that logit and probit fits are pretty similar.
Comparing them to cloglog model, we may say that cloglog model performs better than the others,
generally having smaller residuals in the view of abstract values (norm values).

Next, I show the $\pi_i$ fits verses thecovariate.

```{r}
inv_logit <- function(x, beta_0, beta_1){
    return(exp(beta_0 + beta_1*x) / (1 + exp(beta_0 + beta_1*x)))
}
inv_probit <- function(x, beta_0, beta_1){
    return(pnorm(beta_0 + beta_1*x))
}
inv_cloglog <- function(x, beta_0, beta_1){
    return(1-exp(-exp(beta_0 + beta_1*x)))
}
# grid = seq(1.65, 1.9, 0.005)
plot(log_dose, killed/beetles, xlim=c(1.65, 1.9), cex=3)
curve(inv_logit(x, logit_fit$coefficients[1], logit_fit$coefficients[2]), add=T, col="orange")
curve(inv_probit(x, probit_fit$coefficients[1], probit_fit$coefficients[2]), add=T, col="red")
curve(inv_cloglog(x, cloglog_fit$coefficients[1], cloglog_fit$coefficients[2]), add=T, col="blue")
```

Again, the logit model fit has the orange color, the probit model is red, and cloglog model uses blue.
Black dots are observed ratio, calculated by $y_i/m_i$.

We can see that cloglog fit(blue) is closer to observed ratio for almost all points.
I think that the asymmetry of cloglog link may help to fit better.

### 1-(b)

Please see the separately attached note.

### 1-(c)

Cause all MLE values are given, we can track back fitted $\pi_i$ as well as the model's likelihood value.
Because the equation (1.1) directly gives us the inverse link function, we can use it to get fitted $\pi_i$.

```{r}
#1c: alpha-power logit link
pll_beta0 = -113.625
pll_beta1 = 62.5
pll_alpha = 0.279

pll_inv <- function(x, beta_0, beta_1, alpha){
    nu = beta_0 + beta_1*x
    return(exp(alpha*nu)/((1+exp(nu))^alpha))
}
(pll_fit = pll_inv(log_dose, pll_beta0, pll_beta1, pll_alpha))
```
The above output shows fitted $\pi_i$s at the $i$th data point.

Plus, let me put the $\pi$ fits of the powered-logit link model over the previous plot I showed.

```{r}
plot(log_dose, killed/beetles, xlim=c(1.65, 1.9), cex=3)
curve(inv_logit(x, logit_fit$coefficients[1], logit_fit$coefficients[2]), add=T, col="orange")
curve(inv_probit(x, probit_fit$coefficients[1], probit_fit$coefficients[2]), add=T, col="red")
curve(inv_cloglog(x, cloglog_fit$coefficients[1], cloglog_fit$coefficients[2]), add=T, col="blue")
curve(pll_inv(x, pll_beta0, pll_beta1, pll_alpha), add=T, col="green")
```

The green curve is the new powered-logit link model fit.
In the left tail, it assigns more $pi$ than cloglog fit.
In the middle area (between $1.75$ and $1.80$,) powered-logit fit is giving slightly lower $\pi$ than cloglog fit,
making closer fit to $y/m$.
In the right tail, cloglog and powered logit link fits are very similar.

We can conclude that powered logit fit is quite good, carrying a very flexible asymmetry.
But it is not different very much from cloglog fit, though.
If we want a simpler model, cloglog fit may be better.


```{r}
pll_dev_res <- function(m, y, fitted_pi){
    red_theta = log(fitted_pi/(1-fitted_pi))
    full_theta = log(y/(m-y))
    d = 2*(y*(full_theta - red_theta) + m*log((1+exp(red_theta))/(1+exp(full_theta))))
    sign = ifelse((y/m - fitted_pi)>0, +1, -1)
    return(sign*sqrt(d))
}

#pll dev.resid
pll_dev_res(beetles, killed, pll_fit) #cause m=y for last point, it gives NaN
```

Finally, the last line's output gives us the residual values.
The function named 'pll_dev_res' calculates the deviance residual of the new model.
Since $m_i=y_i$ at $i=8$, my poor, non adjusted funtion yields 'NaN'.
We may use adjusted deviance residual formula for $y=m$ or $y=0$ cases, but let me skip it here.

Other values except for one are quite close to 0.

### 1-(d)

Belows are AIC and BIC values for each model.

```{r}
# 1d. AIC / BIC
pll_log_lik = 0
for(i in 1:8){
    pll_log_lik = pll_log_lik + dbinom(killed[i],beetles[i],pll_fit[i], log=TRUE)
}
# (pll_log_lik)

pll_num_param = 3
(pll_AIC = -2*pll_log_lik + 2*pll_num_param)
(pll_BIC = -2*pll_log_lik + pll_num_param * log(8))

AIC(logit_fit)
BIC(logit_fit)
AIC(probit_fit)
BIC(probit_fit)
AIC(cloglog_fit)
BIC(cloglog_fit)

```

Interestingly, both AIC and BIC select cloglog fit.
It consists of what we found earlier.

Basic logit and probit fit were worse than cloglog or powered logit fit, so they also have higher AIC and BIC values.

Since powered logit model has one more parameter, it gets more penalty from AIC and BIC calculations.
But the fit of powered logit link model was not much improved from cloglog fit, according to the previous observation.
That's why AIC and BIC of powered logit link fit is higher than cloglog fit.