# HW4 problem 3

Seokjun Choi

*Code blocks are R scripts in this problem.*

*Executable stand-alone R script: https://github.com/letsjdosth/GLM/blob/main/hw4q3b.r*

*All scripts are tested in R 4.2.0 64-bit, in Windows 10.*
 

## 3-(b)

```{r}
concentration <- c(0, 62.5, 125, 250, 500)
num_subject <- c(297, 242, 312, 299, 285)
response <- matrix(
    c(15, 1, 281,
    17, 0, 225,
    22, 7, 283,
    38, 59, 202,
    144, 132, 9), 5, 3, byrow=TRUE)
```

Because the model is perfectly decomposed into two binomial GLM models,
we can easily fit the model using the 'glm' function twice.

```{r}
# L1 part fit
L1_y = response[,1]
L1_m_y = response[,2]+response[,3]
L1_fit = glm(cbind(L1_y, L1_m_y)~concentration, family="binomial")
summary(L1_fit)

# L2 part fit
L2_y = response[,2]
L2_m_y = response[,3]
L2_fit = glm(cbind(L2_y, L2_m_y)~concentration, family="binomial")
summary(L2_fit)
```

We can take estimates and standard errors of $\alpha_1, \beta_1$ from the 'L1_fit' summary,
and of $\alpha_2, \beta_2$ from the 'L2_fit' summary.

Below is $\pi$ -plots.
The black curve is for $\pi_1$ (dead), the red is for $\pi_2$ (malformation), and the blue is for $\pi_3$ (normal).

```{r}
grid = seq(0, 601, 1)
inv_logit <- function(x, coeff_vec){
    alpha = coeff_vec[1]
    beta = coeff_vec[2]
    exp_val = exp(alpha+beta*x)
    return(exp_val/(1+exp_val))
}


rho1_on_grid = inv_logit(grid, coef(L1_fit))
rho2_on_grid = inv_logit(grid, coef(L2_fit))
pi1_on_grid = rho1_on_grid
pi2_on_grid = rho2_on_grid * (1 - pi1_on_grid)
pi3_on_grid = 1 - pi1_on_grid - pi2_on_grid
plot(grid, pi1_on_grid, type="l", ylim=c(0,1), main="pi", xlab="concentration", ylab="pi")
points(grid, pi2_on_grid, type="l", col="red")
points(grid, pi3_on_grid, type="l", col="blue")
```

The result is coherent with our intuition. Toxin concentration is associated with malformation or even dead.
(We may not draw some 'causal' interpretation, though. It depends on experimental design, so I need more information to draw it.)
